---
title: "Assignment 2"
author: "Meet Muchhala"
format: html
editor: visual
---use
---

TASK 1: Spatial data visualization (interactive map & choropleth)
For this task, read in the spatial data, and create a professionally formatted and prepared HTML from a Quarto doc in which you: 
## Reading the data and initialization

```{r setup}


library(tidyverse)
library(here)
library(broom)
library(sf)
library(tmap)
library(spatstat)
```

## Loading the datasets 

```{r}

## Setting up the oil spill 2008 events in CA dataset

ca_oilspill_df <- read_csv(here('data', 'oil_spill_ca.csv')) %>% 
  janitor::clean_names()

## Converting Oil Spill coordinates into simple features for ease of use and analysis

ca_oilspill_sf <- ca_oilspill_df %>% 
  drop_na(x, y) %>% 
  st_as_sf(coords = c("x", "y"), crs = 3857) 
  
## Setting up the CA Map with counties boundaries 

ca_map_sf <- read_sf(here('data', 'ca_counties', 'CA_Counties_TIGER2016.shp')) %>% 
  janitor::clean_names() %>% 
  select(name)

```


## Checking and aligning: Coordinate Reference Systems (CRS)

```{r}

st_crs(ca_oilspill_sf) ## <- the CRS system set here is WGS 84/Pseudo-Mercator
st_crs(ca_map_sf) ## <- the CRS system here is WGS 84/Pseudo-Mercator 

st_crs(ca_oilspill_sf) == st_crs(ca_map_sf)
## Hence, no need to transform the features of both the data since they align. 


```

## Interactive map: Oil spill events in California (2008)

```{r}
#| eval: false 
#| echo: true

tmap_mode(mode = "view") 

colors <-  c("Inland" = "brown", "Marine" = "darkblue")



 tm_shape(ca_map_sf) +
  tm_borders() +
  tm_shape(ca_oilspill_sf) +
  tm_dots(col = "inlandmari", shape = 16, palette = colors, title = "Oil Spills Locations") +
   tm_layout(title = "California Oil Spill Incidents 2008")

`````

## Chloreopath map

```{r}
oilspill_counties_sf <- st_join(ca_map_sf, ca_oilspill_sf)

oilspill_counts_sf <- oilspill_counts_sf %>% 
  

```

NOTE: an interactive map using tmap_mode(‘view’) might be very processor intensive and/or create very large files… so if your computer is older and not up to the task, it is OK to create a static map using tmap_mode(‘plot’).
Make a finalized static choropleth map in ggplot in which the fill color for each county depends on the count of inland oil spill events by county for the 2008 oil spill data


